---
title: "Data Science Machine Learning Project"
output: html_document
---

The purpose of this assignment is to predict the class of exercise ("classe") done by participants in a study by examing other measured variables.

First I loaded necessary packages and the data (& test data provided us) into R.
```{r warning=FALSE, message=FALSE}
library(caret)
library(randomForest)
 testing <- read.csv("pml-testing.csv", header = TRUE,na.strings=c("NA",""))
training <- read.csv("pml-training.csv", header = TRUE,na.strings=c("NA",""))
```
I removed any obviously uninformative columns as well as those with incomplete data ('NA' values).
```{r}
removecols <- grepl("X|user_name|new_window|timestamp|num_window", colnames(testing))
testing <- testing[, !removecols]
training <- training[, !removecols]
training <- training[ , ! apply( training , 2 , function(x) any(is.na(x)) ) ]
testing <- testing[ , ! apply( testing , 2 , function(x) any(is.na(x)) ) ]
```
In order to prepare for cross validation I partitioned part of the training set and used the partition to create two separate datasets.  I chose a 60/40 ratio.
```{r}
partitioning = createDataPartition(y=training$classe, p=0.6, list=FALSE)
partValid = training[-partitioning, ]
partTraining = training[partitioning, ]
```
Next, I fit a model to our partioned training data using random forest classification. (Note that I did not use formula for this so as to cut down on computing time.) 

That fit model was then cross validated against both the larger training data set and the smaller validation subset to check that its accuracy/error rate was acceptable.
```{r}
modFit <- randomForest(partTraining[ ,1:52], partTraining$classe)
crossTrain <- predict(modFit, partTraining[ ,1:52])
crossValidate <- predict(modFit, partValid[ ,1:52])
accuracy <- postResample(partValid$classe, crossValidate)
accuracy <- accuracy[[1]]
```
Accuracy:
```{r}
print(accuracy)
```
Out of Sample error rate:
```{r}
print((1-accuracy))
```
With high accuracy and a low error rate, I then went ahead and used the fit model to predict for the provided test data:
```{r}
predict <- predict(modFit, newdata=testing)
print(predict)
```


